Sampbias, a method to evaluate geographic sampling bias in species distribution data

### 2019-11-29 manuscript.rmd
gitignore and bibliogrphy file

Geographic sampling bias, the fact that sampling effort is spatially biased, rather than equally distributed over a given study area is prevalent in all non-systematically collected datasets of species distributions. Many factors can affect sampling effort, such as socio-economic [i.e. national research spending, history of scientific research, @Meyer2016] and political factors [armed conflict, democratic rights, @Ryden2019] or physical accessibility [i.e. distance to a road or river, terrain conditions, slope, @Yang2014; @Botts2011]. Especially physical accessibility is omnipresent as a biasing factor [e.g. @Lin2015; @Engemann2015], across spatial scales, and the term "roadside bias has been coined for it. In practice, this means that most species observations (occurrence points) are made in or near cities, along roads and rivers, and near other human settlements (such as airports). Less observations come from the middle of a tropical rainforest or from a mountain top. Interestingly, since the observation of different taxonomic groups has different challenges, geographic sampling bias and the effect of accessibility may differ among taxonomic groups [@Vale2012].

@@

# Author contributions
All authors conceived of this study, AZ and DS developed the statistical algorithm, AZ wrote the R-package and Shiny app, AZ and DS wrote the manuscript with contributions from AA.

>>>>

Geographic sampling bias, the fact that sampling effort is spatially biased, rather than equally distributed over a given study area is prevalent in all non-systematically collected datasets of species distributions. Many factors can affect sampling effort, such as socio-economic [i.e. national research spending, history of scientific research, @Meyer2015, @Daru2018] and political factors [armed conflict, democratic rights, @Ryden2019] or physical accessibility [i.e. distance to a road or river, terrain conditions, slope, @Yang2014; @Botts2011]. Especially physical accessibility is omnipresent as a biasing factor [e.g. @Lin2015; @Engemann2015], across spatial scales, and the term "roadside bias has been coined for it. In practice, this means that most species observations (occurrence points) are made in or near cities, along roads and rivers, and near other human settlements (such as airports). Less observations come from the middle of a tropical rainforest or from a mountain top. Interestingly, since the observation of different taxonomic groups has different challenges, geographic sampling bias and the effect of accessibility may differ among taxonomic groups [@Vale2012].

@@

# Author contributions
All authors conceived of this study, AZ and DS developed the statistical algorithm, AZ and DS wrote the R-package and AZ the Shiny app, AZ and DS wrote the manuscript with contributions from AA.
